TacticToe: Learning the Game of Tactics

### 2018-01-30 tactictoe_jv.tex
minor

\begin{itemize}
\item Proof recording at the tactic level is made more precise. Pattern 
matching constructions, opening of submodules are supported. The different 
recording steps (parsing, unfolding and replaying) are regrouped into a single 
\sml function.
\item MCTS (Monte Carlo tree search) now replaces A* in our proof search 
algorithm. MCTS gives a more balanced feedback to guide proof search by 
comparing subtrees instead of leafs. The policy and evaluation are learned 
through supervised learning.
\item Proof guidance required by the MCTS algorithm is given by predictors for 
three kind of objects: tactics, theorems and list of goals.
\item The orthogonalization process that eliminates redundant tactics is 
formally described.
\item A tactic abstraction method is introduced. It enable us to create 
more general and flexible tactics where tactic arguments are dynamically 
predicted.
\item The internal ATP \metis is complemented 
by asynchronous calls to \eprover to help \tactictoe during proof search.
\todoi{More on evaluation??}
\item Evaluation of \tactictoe success rate relative to 
the length of the original proofs.
\item User interaction is enhanced by the minimization and prettification 
applied on the generated proof.
\end{itemize}

@@

Descendants of $a$ through $g$ by:
\[Descendant(a,g) =_{def} 
  \lbrace a \rbrace \cup \bigcup_{a' \in Children(a,g)} Descendant(a')\]

@@

The current evaluation $CurEvaluation(a_i)$ is the average evaluation of 
all descendants of $a_i$, through its open goal $g_i$, including node extension 
failures:

\[CurEvaluation(a_i) = 
  \sum_{a' \in Descendant(a_i,g_i)} \frac{PriorEvaluation(a')} {card\ 
  Descendant(a_i,g_i) + Failure(a_i)}\]

@@

\subsection{Increasing \tactictoe'sknowledge}\label{sec:perfect_exp}

@@

\begin{table}[h]
\centering\ra{1.3}
\small
\begin{tabular}{lll}
\toprule
 & Additional features & Solved \\
\midrule
$\emptyset$   & none & 119 (43.6\%) \\ %NONEv2
$\mathbb{F}_1$& tactic prediction & 121 (44.3\%) \\ %AFTERTACv2
$\mathbb{F}_2$& token recombination  & 137 (50.2\%) \\ %AFTERTOKENv2
$\mathbb{F}_3$& theorem prediction  & 207 (75.8\%) \\ %AFTERTHMTHMv2
$\mathbb{F}_4$& term prediction & 223 (81.7\%) \\ %AFTERALLv2
$\mathbb{H}$  & all & 239 (87.5\%) \\ %AFTERSMALLv2
% $AFTERv2$       & $\mathfrak{H}$ & 245 (89.7\%)  \\
\bottomrule
\end{tabular}
\caption{\label{tab:featue_param} Effect of augmenting \tactictoe's knowledge 
with tactics from the proof of the tested theorem. The experiments was run with 
a time out of 5sec on an older version of \tactictoe, as can be seen in the low 
success rate of the
default strategy.} 
\end{table}

And we observe in Table~\ref{tab:featue_param} a 
slight increase compared to the default \tactictoe. We then included all 
vectors that contains tactics that can be rebuild from tokens of
tactics in our database. 
Our search algorithm proved two proofs less that if it had perfect prediction 
so it means that our prediction algorithm (policy) is doing a good  job at 
least in this early settings where proofs are easy. 
Suprisingly, being able to recombine tokens of tactics from the database is 
barely sufficient to prove more than half of the theorems in this dataset. The 
Therefore, additionally including in tactics theorems and terms that did not 
appear in previous tactics is essential. Creating from scratch other arguments 
may not help as much. This is why we restricted ourselves to predicting 
arguments of theorem list type and the term type during tactic synthesis. 
Finally our proof search algorithm sometimes fails even if it knows all the 
tactic necessary. The two main reasons are that either a crucial tactic 
exceeds its timeout or that the human proof is too deep to be rediscovered 
(more than 10 steps long).In all those experiments orthogonalization is 
performed with 
radius 20.
%
%\subsubsection{Orthogonalization Abstraction}
%Here, we test the benefit of different kind of tactic abstraction.
%All experiments in this table are using orthogonalization.
%
%\begin{table}[t]
%\centering\ra{1.3}
%\small
%\begin{tabular}{ll}
%\toprule
% Technique & Solved \\
%\midrule
%% holyhammer & 79 \\
%  & 122\\
% Orthogonalization & 129\\
%% ortho + term & 126\\    
% Orthogonalization + Abstraction & 160\\
%% ortho + \metis & 162\\
% %ortho + metis + thmortho & 155\\
% %ortho + metis + stacortho & 160\\
%% ortho + \metis + theorem list & 175\\ 
%\bottomrule
%\end{tabular}
%\caption{\label{tab:cfot_param} 
% Effect of orthogonalization and abstraction}
%\end{table}
%We want to know what the optimal number of predictions for tactics different 
%from \metis since \metis has a special status in our algorithm is handled 
%independently.

%\subsubsection{ATP integration}
%
%
%\todo{paragraph is copied}
%The best setting for \metis was to be in previous experiments to be 16
%theorems an The best strategy which does not rely on the ``small hammer'' 
%approach $D_9$ 
%will be called \tactictoe(NH) (for no ``small hammer'') in the remaining part 
%of the paper, and the best strategy relying on the approach, $D_{19}$, will be 
%referred to as \tactictoe(SH) (``small hammer'').
%
%\begin{table}[h!]
%\centering\ra{1.3}
%\small
%\begin{tabular}{ll}
%\toprule
%  & Solved \\
%\midrule
%   \eprover &  106 \\
%   \tactictoe &  139 \\
%   \tactictoe + \metis & 182 \\
%   \tactictoe + \eprover & 182 \\
%   %x tactictoe + async holyhammer (10 thread) & 183 \\
%   \tactictoe + \metis + \eprover & 190 \\
%  
%\bottomrule
%\end{tabular}
%\caption{Experiments with a timeout of 60 seconds \label{tab:param}}
%\end{table}
%
%Slight error in this experiments were eprover predictions have been made from 
%the preselected 500 theorems instead of the full library. But even with the 
%added overhead, we already see 
%the benefit of using stronger external prover than \metis.

@@

\begin{table}[h!]
\centering\ra{1.3}
\small
\begin{tabular}{llc}
\toprule
  & Parameters & Solved \\
\midrule
   \eprover   & 128 premises & 2472 (34.5\%)\\ 
   \tactictoe & best parameters without \eprover & 4741 (66.2\%)\\
   combined   &              & 4946 (69.0\%)\\
\bottomrule
\end{tabular}
\caption{Evaluation on 7164 top-level theorems of the \holfour standard library 
\label{tab:_param}}
\end{table}

>>>>

\begin{itemize}
\item Proof recording at the tactic level is made more precise. Pattern 
matching constructions, opening of submodules are supported. The different 
recording steps (parsing, unfolding and replaying) are regrouped into a single 
\sml function.
\item Monte Carlo tree search~~\cite{montecarlo} (MCTS) replaces A* as our 
search algorithm. The MCTS algorithm gives a more balanced feedback on 
the success of 
each proof step by
comparing subtrees of the search tree instead of leafs. The policy and 
evaluation are learned 
through supervised learning.
\item Proof guidance required by MCTS is given by predictors for 
three kind of objects: tactics, theorems and list of goals.
\item The orthogonalization process that eliminates redundant tactics is 
formally described.
\item A tactic abstraction method is introduced. It enable us to create 
more general and flexible tactics where tactic arguments are dynamically 
predicted.
\item The internal ATP \metis is complemented 
by asynchronous calls to \eprover to help \tactictoe during proof search.
\todoi{More on evaluation??}
\item Evaluation of \tactictoe success rate relative to 
the length of the original proofs.
\item User interaction is enhanced by the minimization and prettification 
applied on the generated proof.
\end{itemize}

@@

%Descendants of $a$ through $g$ by:
%\[Descendant(a,g) =_{def} 
%  \lbrace a \rbrace \cup \bigcup_{a' \in Children(a,g)} Descendant(a')\]

@@

The current evaluation $CurEvaluation(a_i)$ is the average evaluation of 
all descendants of $a_i$ including node extension 
failures:

\[CurEvaluation(a_i) = 
  \sum_{a' \in Descendant(a_i)} \frac{PriorEvaluation(a')} {card\ 
  Descendant(a_i) + Failure(a_i)}\]

@@

\subsection{Increasing \tactictoe's knowledge}\label{sec:perfect_exp}

@@

\begin{table}[h]
\centering\ra{1.3}
\small
\begin{tabular}{lll}
\toprule
 & Additional features & Solved \\
\midrule
$\emptyset$   & none & 119 (43.6\%) \\ %NONEv2
$\mathbb{F}_1$& tactic prediction & 121 (44.3\%) \\ %AFTERTACv2
$\mathbb{F}_2$& token recombination  & 137 (50.2\%) \\ %AFTERTOKENv2
$\mathbb{F}_3$& theorem prediction  & 207 (75.8\%) \\ %AFTERTHMTHMv2
$\mathbb{F}_4$& term prediction & 223 (81.7\%) \\ %AFTERALLv2
$\mathbb{H}$  & all & 239 (87.5\%) \\ %AFTERSMALLv2
% $AFTERv2$       & $\mathfrak{H}$ & 245 (89.7\%)  \\
\bottomrule
\end{tabular}
\caption{\label{tab:featue_param} Effect of augmenting \tactictoe's knowledge 
with tactics from the proof of the tested theorem. The experiments was run with 
a time out of 5 seconds on an older version of \tactictoe, as can be seen in 
the low 
success rate of the
default strategy.} 
\end{table}

And we observe in Table~\ref{tab:featue_param} a 
slight increase compared to the default \tactictoe. We then included all 
vectors that contains tactics that can be rebuild from tokens of
tactics in our database. 
Our search algorithm proved two proofs less that if it had perfect prediction 
so it means that our prediction algorithm (policy) is doing a good  job at 
least in this early settings where proofs are easy. 
Suprisingly, being able to recombine tokens of tactics from the database is 
barely sufficient to prove more than half of the theorems in this dataset. The 
Therefore, additionally including in tactics theorems and terms that did not 
appear in previous tactics is essential. Creating from scratch other arguments 
may not help as much. This is why we restricted ourselves to predicting 
arguments of theorem list type and the term type during tactic synthesis. 
Finally our proof search algorithm sometimes fails even if it knows all the 
tactic necessary. The two main reasons are that either a crucial tactic 
exceeds its timeout or that the human proof is too deep to be rediscovered 
(more than 10 steps long).In all those experiments orthogonalization is 
performed with 
radius 20.

@@

\begin{table}[h!]
\centering\ra{1.3}
\small
\begin{tabular}{llc}
\toprule
  & Parameters & Solved \\
\midrule
   \eprover   & & 2472 (34.5\%)\\ 
   \tactictoe & best parameters without \eprover & 4741 (66.2\%)\\
\midrule
   Total  &              & 4946 (69.0\%)\\
\bottomrule
\end{tabular}
\caption{Evaluation on 7164 top-level theorems of the \holfour standard library 
\label{tab:_param}}
\end{table}
